---
title: "DABI_25_MA7419_CW2"
output: ioslides_presentation
---

## Road trafic accidents resulting in a casualty
  - Question 1: Analyze accidents severity in UK for the last 5 years
  - Visualization: Slides 4 - 10
  - Question 2: Compare accident deaths for UK and European Union in the last decade
  - Visualization: Slides 11 - 13

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, warning=FALSE, message=FALSE}

library(tidyverse)
library(dplyr)
library(latexpdf)
library(knitr)
library(readxl)
library(ggplot2)
library(janitor)
library(RgoogleMaps)
library(stats19)
library(wordcloud2)
library(rworldmap)
library(ranger)
```


```{r, warning=FALSE, message=FALSE}

casualty_data <- read_csv("/home/dd231/R/MA7419/FDS/casualty_five_years.csv")
accident_data <-read_csv("/home/dd231/R/MA7419/FDS/accident_five_years.csv")
```

```{r}

# Total Casualties in UK 2016-2021

casualty_total_year_wise<-accident_data  %>%
  group_by(accident_year) %>%
  summarise(number_of_casualties = sum(number_of_casualties))

# Total Accidents in UK 2016-2021

accident_total_year_wise<-accident_data  %>%
  group_by(accident_year) %>%
  summarise(n=n())
names(accident_total_year_wise)[names(accident_total_year_wise) == "n"] <- "number_of_accidents"
```

## Contents of the presentation
- Accidents & Casualities in UK between 2016 - 2020
- Categories of Accident Severity
- Accident Serverity in different Districts
- Geospatial mapping on Brent(sample): Based on accident severity
- Accident severity based on age band
- Listing the importance variables of accident severity (Random Forest)
- Listing the importance variables of Casualty severity (Random Forest)
- Comparing total accident deaths:UK Vs European Countries
- EU Countries with more than 60% Accident Deaths per million inhabitants in 2010 & 2020


## Accidents & Casualties in UK between 2016 - 2020
```{r, warning=FALSE, message=FALSE}

#Joining accidents & casualties data

year<-accident_total_year_wise$accident_year
casualty<-casualty_total_year_wise$number_of_casualties
accident<-accident_total_year_wise$number_of_accidents
accident_and_casualty_total<-full_join(casualty_total_year_wise, accident_total_year_wise)

#plotting accidents and casualties for the last five years

ggplot()+
  geom_line(data=accident_and_casualty_total,aes(y=accident,x= year,colour="Accidents"),size=1 )+
  geom_line(data=accident_and_casualty_total,aes(y=casualty,x= year,colour="Casualties"),size=1) +
  scale_color_manual(name = "Accidents & Casualities", values = c("Accidents" = "blue", "Casualties" = "red"))+labs(title="
            Casualties and Accidents in UK 2016-2020",
x ="Year", y = "Total Frequency")+ theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"),
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white"))

```


```{r, warning=FALSE, message=FALSE}

#Description of the casualty severity

casualty_severity_total_year_wise<-tabyl(casualty_data, accident_year, casualty_severity)
names(casualty_severity_total_year_wise)[names(casualty_severity_total_year_wise) == "1"] <- "fatality"
names(casualty_severity_total_year_wise)[names(casualty_severity_total_year_wise) == "2"] <- "serious_injury"
names(casualty_severity_total_year_wise)[names(casualty_severity_total_year_wise) == "3"] <- "slight_injury"
```


## Categories of Accident Severity
```{r, warning=FALSE, message=FALSE}

# Types of Accident severity

accident_year<-casualty_severity_total_year_wise$accident_year
fatality<-casualty_severity_total_year_wise$fatality
serious_injury<-casualty_severity_total_year_wise$serious_injury
slight_injry<-casualty_severity_total_year_wise$slight_injury


#plotting accident severity category-wise for the last five years

ggplot()+
  geom_line(data=casualty_severity_total_year_wise,aes(y=fatality,x= accident_year,colour="Fatalities"),size=1 )+
  geom_line(data=casualty_severity_total_year_wise,aes(y=serious_injury,x= accident_year,colour="Serious Injuries"),size=1) +
  geom_line(data=casualty_severity_total_year_wise,aes(y=slight_injry,x= accident_year,colour="Slight Injuries"),size=1) +
  scale_color_manual(name = "Categories of Accident Severity", values = c("Fatalities" = "red", "Serious Injuries" = "black","Slight Injuries" = "purple"))+labs(title="
            Accident Severity Types in UK 2016-2020",
x ="Year", y = "Number of Injuries")+ theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"),
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white"))
```


```{r warning=FALSE, message=FALSE}

#Extracting stats19 data with description for casualty year-wise

casualty_year_2016 = get_stats19(year = 2016, type = "casualty")
casualty_year_2017 = get_stats19(year = 2017, type = "casualty")
casualty_year_2018 = get_stats19(year = 2018, type = "casualty")
casualty_year_2019 = get_stats19(year = 2019, type = "casualty")
casualty_year_2020 = get_stats19(year = 2020, type = "casualty")

cas_joined_all<-full_join(casualty_year_2016, casualty_year_2017)
cas_joined_all<-full_join(cas_joined_all, casualty_year_2018)
cas_joined_all<-full_join(cas_joined_all, casualty_year_2019)
cas_joined_all<-full_join(cas_joined_all, casualty_year_2020)

#Extracting stats19 data with description for accident year-wise

accident_year_16 = get_stats19(year = 2016, type = "accident")
accident_year_17 = get_stats19(year = 2017, type = "accident")
accident_year_18 = get_stats19(year = 2018, type = "accident")
accident_year_19 = get_stats19(year = 2019, type = "accident")
accident_year_20 = get_stats19(year = 2020, type = "accident")
accident_year_17$longitude<-as.numeric(accident_year_17$longitude)
accident_year_17$latitude<-as.numeric(accident_year_17$latitude)
total_accident_joined<-full_join(accident_year_16, accident_year_17)

total_accident_joined<-full_join(total_accident_joined, accident_year_18)
total_accident_joined<-full_join(total_accident_joined, accident_year_19)
total_accident_joined<-full_join(total_accident_joined, accident_year_20)
total_accident_joined<-na.omit(total_accident_joined)
 
```

## Accident Severity in different Districts
```{r}

# Finding the frequency of each local authority districts for word cloud

total<-total_accident_joined %>%count(local_authority_district, sort = TRUE, name = "freq")

# Finding the local authority districts with high accident severity

total_accident_joined %>%
  count(local_authority_district, sort = TRUE, name = "freq") %>%
  filter(freq > 10) %>%
  wordcloud2()
```


## Geospatial mapping on Brent(sample): Based on Accident Severity
```{r warning=FALSE, message=FALSE}

#Mapping the casualty severity data on static google map

accident_data<-na.omit(accident_data)
casualty_class_driver_rider<-subset(accident_data,local_authority_district == 28)
casualty_class_driver_rider_fatal <- subset(casualty_class_driver_rider,accident_severity == 2)
longitude <- c(casualty_class_driver_rider$longitude)
latitude <- c(casualty_class_driver_rider$latitude)
center = c(mean(latitude),mean(longitude))
col=as.numeric(casualty_class_driver_rider$accident_severity)
zoom <-12
terrmap <- GetMap(center = center, zoom = zoom,maptype= "mapmaker-hybrid", 
                   destfile = tempfile("mapmaker-hybrid", fileext = ".png"),markers=markers,verbose=0)

PlotOnStaticMap(terrmap,lat=latitude,lon=longitude,col=c('red','black','purple'),pch=15)
unique_col <- unique(col)
unique_col_fatal <- replace(unique_col,1,"Fatal")
unique_col_serious <- replace(unique_col_fatal,2,"Serious")
unique_col_slight <- replace(unique_col_serious,3,"Slight")

legend("topleft", legend = unique_col_slight, fill = c('red','black','purple'),
bg = rgb(0.1,0.1,0.1,0.3))
```




```{r, warning=FALSE, message=FALSE}

# Removing the null values from the accident and casualty data after joining

total_casualty_accident_data<-full_join(accident_data, casualty_data)

total_casualty_accident_data<-na.omit(total_casualty_accident_data)
```

## Accident Severity based on Age-band
```{r}

# Mosaic plot of accident severity on the basis of age band

mosaicplot(age_band_of_casualty~accident_severity,data=total_casualty_accident_data,col=c("red","black","purple"),xlab="Age Band",ylab="Accident Severity",main = "Accident Severity: Based on Age Band")

```
----- 1 : 0-5 age, 2 : 6-10 age, 3 : 11-15 age, 4 : 16-20 age, 5 : 21-25 age, 6 : 26-35 age, 7 : 36-45 age, 8 : 46-54 age, 9 : 55-64 age, 10 : 65-74, 11 : 75 plus age



```{r}

# Referred from : https://stackoverflow.com/questions/56304698/how-do-i-plot-the-variable-importance-of-my-trained-rpart-decision-tree-model

# Feature extractions of accident severity using Random Forest Algorithm

accident_data<-subset(accident_data, select = -c(1,3:5) )

model_rf <-ranger(
  accident_severity ~.,
  data = accident_data,
  importance = "impurity",
  num.trees = 500)

variable_importance_data <- data.frame(imp = model_rf$variable.importance)
```

## Listing the importance variables of Accident Severity (Random Forest)
```{r}

# Plotting the importance variables for accident severity

listing_variable_importance_data <- variable_importance_data %>%
  tibble::rownames_to_column() %>%
  dplyr::rename("variable" = rowname) %>%
  dplyr::arrange(imp) %>%
  dplyr::mutate(variable = forcats::fct_inorder(variable))
ggplot2::ggplot(listing_variable_importance_data) +
  geom_col(aes(x = variable, y = imp),
           col = "black", show.legend = F) +
  coord_flip() +
  scale_fill_grey() +
  theme_bw()
```

```{r}

# Referred from : https://stackoverflow.com/questions/56304698/how-do-i-plot-the-variable-importance-of-my-trained-rpart-decision-tree-model  

# Feature extractions of casualty severity using Random Forest Algorithm

casualty_data<-subset(casualty_data, select = -c(1,3:5) )

model_rf <-ranger(
  casualty_severity ~.,
  data = casualty_data,
  importance = "impurity",
  num.trees = 500)

imp_variable_data <- data.frame(imp = model_rf$variable.importance)
```


## Listing the importance variables of Casualty Severity (Random Forest)
```{r}

# Plotting the importance variables for casualty severity

listing_imp_variable_data <- imp_variable_data %>%
  tibble::rownames_to_column() %>%
  dplyr::rename("variable" = rowname) %>%
  dplyr::arrange(imp) %>%
  dplyr::mutate(variable = forcats::fct_inorder(variable))
ggplot2::ggplot(listing_imp_variable_data) +
  geom_col(aes(x = variable, y = imp),
           col = "black", show.legend = F) +
  coord_flip() +
  scale_fill_grey() +
  theme_bw()
```

```{r}

# Detailed view of casualty type on the basis of accident severity

boxplot(casualty_severity~casualty_type,data=casualty_data,ylab= "casualty severity", xlab= "casualty_type", main= "casualty severity on the basis of casualty type")
```


```{r}

# European Union data of accident deaths in the last decade

eu_data <- read_excel("/home/nrs14/Nag/FDS_PROJECT/eudata.xls")

```


```{r}

# finding the total number of accidents for each country in EU

eu_data_melt<- reshape2::melt(eu_data, id = c("country"))
colnames(eu_data_melt)[2] <-"year"
colnames(eu_data_melt)[3] <-"no_of_accidents"
eu_data_melt<- eu_data_melt %>% group_by(country) %>% summarise_at(vars(no_of_accidents),list(no_of_accidents=sum))
eu_data_melt <- eu_data_melt[order(-eu_data_melt$no_of_accidents),]
eu_data_melt <- eu_data_melt %>% slice(-c(1))

```


```{r}

# selecting the EU countries with over 10000 deaths in the last decade

eu_data_greater_thousand <- eu_data_melt %>% filter(no_of_accidents >= 10000)
 
```


## Comparing total Accident Deaths:UK Vs European Countries
```{r}

# Visualizing accident deaths for selected EU countries

eu_data_greater_thousand$country[eu_data_greater_thousand$country == "Germany (until 1990 former territory of the FRG)"]<-"Germany"
eu_data_greater_thousand$country[eu_data_greater_thousand$country == "United Kingdom"]<-"UK"

eu_data_greater_thousand_chart<- ggplot(eu_data_greater_thousand, aes(x=country, y=no_of_accidents, fill=country))+
geom_bar(width = 1, stat = "identity")
eu_data_greater_thousand_chart

```


```{r, warning=FALSE, message=FALSE}

# European Union data on accident deaths for PIN (Performance Index: Road safety) countries

data_europin_inhabitants <- read_excel("/home/nrs14/Nag/FDS_PROJECT/eu_pin_countries.xlsx", sheet=2)

```



```{r, warning=FALSE, message=FALSE}

#Referred from: https://slcladal.github.io/maps.html

# Selecting accident deaths per million inhabitants (in %) for EU PIN countries

data_europin_inh_top_2010<-subset(data_europin_inhabitants,select = c(country,Deaths_per_mln_inhabitants_2010))
data_europin_inh_top_2020<-subset(data_europin_inhabitants,select = c(country,Deaths_per_mln_inhabitants_2020))

# Filtering accident deaths per million inhabitants (more than 60%) for EU PIN countries

data_europin_inh_top_sel_2020<- data_europin_inh_top_2020 %>% filter(Deaths_per_mln_inhabitants_2020 > 60)
data_europin_inh_top_sel_2010<- data_europin_inh_top_2010 %>% filter(Deaths_per_mln_inhabitants_2010 > 60)

data_europin_inh_top_2010

```


```{r, warning=FALSE, message=FALSE}

# Matching country code with world map (2010)

visitedMap <- joinCountryData2Map(data_europin_inh_top_sel_2010, 
                                  joinCode = "ISO2",
                                  nameJoinColumn = "country")
```

## EU Countries with more than 60% Accident Deaths per million inhabitants in 2010
```{r, warning=FALSE, message=FALSE}

#plotting the accident deaths on Europe map for 2010

mapParams <- mapCountryData(visitedMap, 
                            nameColumnToPlot="Deaths_per_mln_inhabitants_2010", 
                            oceanCol = "azure2",
                            catMethod = "categorical",
                            missingCountryCol = gray(.8),
                            mapRegion = "Europe",
                            colourPalette = c("coral",
                                              "coral2",
                                              "coral3", "orangered", 
                                              "orangered3", "orangered4"),
                            addLegend = F,
                            mapTitle = "Deaths per million inhabitants 2010",
                            border = NA)

do.call(addMapLegendBoxes, c(mapParams,
                             x = 'right',bg = "transparent",bty="n"))
```



```{r, warning=FALSE, message=FALSE}

# Matching country code with world map (2020)

visitedMap <- joinCountryData2Map(data_europin_inh_top_sel_2020, 
                                  joinCode = "ISO2",
                                  nameJoinColumn = "country")
```

## EU Countries with more than 60% Accident Deaths per million inhabitants in 2020
```{r, warning=FALSE, message=FALSE}

#plotting the accident deaths on Europe map for 2020

mapParams <- mapCountryData(visitedMap, 
                            nameColumnToPlot="Deaths_per_mln_inhabitants_2020", 
                            oceanCol = "azure2",
                            catMethod = "categorical",
                            missingCountryCol = gray(.8),
                            mapRegion = "Europe",
                            colourPalette = c("coral",
                                              "coral2",
                                              "coral3", "orangered", 
                                              "orangered3", "orangered4"),
                            addLegend = F,
                            mapTitle = "Deaths per million inhabitants 2020",
                            border = NA)
# add legend and display map
do.call(addMapLegendBoxes, c(mapParams,x = 'right',bg = "transparent",bty="n"))
```



```{r}
citation("tidyverse")
citation("dplyr")
citation("latexpdf")
citation("knitr")
citation("readxl")
citation("ggplot2")
citation("janitor")
citation("RgoogleMaps")
citation("stats19")
citation("wordcloud2")
citation("rworldmap")
citation("ranger")
```

